---
title: "Project"
subtitle: EU frameworks FP1 - FP7 and Horizon2020 funded research projects
output:
  html_document: 
    toc: yes
  html_notebook: default
---

# Introduction
Here we analyse a dataset containing European Union (EU) funded research projects under frameworks

* [FP1](https://data.europa.eu/euodp/en/data/dataset/fp1-cordis) (1984-1987),
* [FP2](https://data.europa.eu/euodp/en/data/dataset/cordis-fp2) (1987-1991),
* [FP3](https://data.europa.eu/euodp/en/data/dataset/cordis-fp3) (1990-1994),
* [FP4](https://data.europa.eu/euodp/en/data/dataset/cordisfp4projects) (1994-1998),
* [FP5](https://data.europa.eu/euodp/en/data/dataset/cordisfp5projects) (1998-2002),
* [FP6](https://data.europa.eu/euodp/en/data/dataset/cordisfp6projects) (2002-2006),
* [FP7](https://data.europa.eu/euodp/en/data/dataset/cordisfp7projects) (2007-2013), 
* [Horizon2020](https://data.europa.eu/euodp/en/data/dataset/cordisH2020projects) (2014-2020).

Coordinates for countries are taken [here](https://developers.google.com/public-data/docs/canonical/countries_csv). 

The subject (SIC) codes are taken [here](https://data.europa.eu/euodp/en/data/dataset/cordisref-data)

# Data import
```{r, error=FALSE, message=FALSE, warning=FALSE}
#Load some libraries
library(tidyverse) #Data handling, plotting
library(stringr) #String manipulation
library(tm) #Text mining functionality
#library(plotly) #Interactive ggplots
library(tidytext) #Text manipulation
library(SnowballC) #Word stemming
#library(widyr) #Pairwise correlations etc
library(maps) #Drawing maps
library(geosphere) #Computing great cricle lines
library(igraph) #Drawing graphs
library(ggraph) #Drawing graphs
#library(irlba) #Truncated svd on large/sparse matrices
library(wordcloud) #Wordclouds
#library(lvplot)
library(lubridate)
library(magrittr)
```

Read in the datasets using `read_csv2` because the files are `;`-delimited. This is apparently common in Europe because `,` is used as decimal delimiter. Each dataset needs some custom modifications when importing but by using readr we can get them to common shape with little trouble

```{r, message=FALSE}
fp1 <- read_csv2("../dat/eu/fp1.csv", 
                 col_types = cols(startDate = col_date(format = "%d/%m/%Y"),
                                  endDate = col_date(format = "%d/%m/%Y"),
                                  totalCost = col_integer(),
                                  ecMaxContribution = col_integer())) %>%
  select(rcn, status, title, startDate, endDate, objective, 
         cost = totalCost, contr = ecMaxContribution, fundingScheme, coordinator, 
         coordinatorCountry, participants, participantCountries, subjects)

fp2 <- read_csv2("../dat/eu/fp2.csv") %>% 
  select(rcn, status, title, startDate, endDate, objective, 
         cost = totalCost, contr = ecMaxContribution, fundingScheme, coordinator, 
         coordinatorCountry, participants, participantCountries, subjects)

fp3 <- read_csv2("../dat/eu/fp3.csv") %>% 
  select(rcn, status, title, startDate, endDate, objective, 
         cost = totalCost, contr = ecMaxContribution, fundingScheme, coordinator, 
         coordinatorCountry, participants, participantCountries, subjects)

fp4 <- read_csv2("../dat/eu/fp4.csv", 
                 col_types = cols(
                   `Start Date` = col_date(format = "%d/%m/%Y"),
                   `End Date` = col_date(format = "%d/%m/%Y"))
                 ) %>%
  select(rcn = RCN, status = Status, title = `Project Title`, 
         startDate = `Start Date`, endDate = `End Date`, objective = Objectives, 
         cost = `Total Cost`, contr = `Total Funding`, fundingScheme = `Contract Type`, 
         coordinatorCountry = `Coordinator Country`, 
         participantCountries = `Contractor Country`, subjects = Subject) %>%
  mutate(coordinator = parse_character(NA), participants = parse_character(NA))

fp5 <- read_csv2("../dat/eu/fp5.csv") %>%
  select(rcn, status, title, startDate, endDate, objective, 
         cost = totalCost, contr = ecMaxContribution, fundingScheme, coordinator, 
         coordinatorCountry, participants, participantCountries, subjects)

fp6   <- read_csv2("../dat/eu/fp6.csv") %>%
  select(rcn, status, title, startDate, endDate, objective, 
         cost = totalCost, contr = ecMaxContribution, fundingScheme, coordinator, 
         coordinatorCountry, participants, participantCountries, subjects)

fp7   <- read_csv2("../dat/eu/fp7.csv") %>%
  select(rcn, status, title, startDate, endDate, objective, 
         cost = totalCost, contr = ecMaxContribution, fundingScheme, coordinator, 
         coordinatorCountry, participants, participantCountries, subjects)

h2020 <- read_csv2("../dat/eu/h2020.csv") %>%
  select(rcn, status, title, startDate, endDate, objective, 
         cost = totalCost, contr = ecMaxContribution, fundingScheme, coordinator, 
         coordinatorCountry, participants, participantCountries, subjects)

projects <- bind_rows(fp1, fp2, fp3, fp4, fp5, fp6, fp7, h2020) %>%
  arrange(startDate) %>%
  mutate(year = year(startDate),
         ratio = contr/cost)

#Clean up
rm(fp1, fp2, fp3, fp4, fp5, fp6, fp7, h2020)

#Change some EU codes into ISO codes
projects <- projects %>%
  mutate(coordinatorCountry = str_replace_all(coordinatorCountry, "UK", "GB")) %>%  #UK
  mutate(coordinatorCountry = str_replace_all(coordinatorCountry, "EL", "GR")) %>% #Greece
  mutate(coordinatorCountry = str_replace_all(coordinatorCountry, "KO", "XK")) %>% #Kosovo
  mutate(coordinatorCountry = str_replace_all(coordinatorCountry, "CS", "RS")) %>%#Serbia
  mutate(participantCountries = str_replace_all(participantCountries, "UK", "GB")) %>%  #UK
  mutate(participantCountries = str_replace_all(participantCountries, "EL", "GR")) %>% #Greece
  mutate(participantCountries = str_replace_all(participantCountries, "KO", "XK")) %>% #Kosovo
  mutate(participantCountries = str_replace_all(participantCountries, "CS", "RS"))

#Country coordinates
# na="" necessary, otherwise country code for Namibia (NA) is interpreted as <NA>
countries <- read_tsv("../dat/eu/countries.csv", na = "")

#subjects sic codes
sic <- read_csv2("../dat/eu/sic_codes.csv", na = "") %>% 
  filter(language == "en") %>%
  select(-language) %>%
  mutate(title = tolower(title),
         code = tolower(code))

```

Glimpse the dataset
```{r}
projects %>% glimpse
```

#Data cleaning, imputation
Let's look at the individual variables

* rcn
```{r}
#--- rcn ---
#All projects have unique id
projects %$% any(is.na(rcn)) #FALSE
projects %>% count(rcn) %>% filter(n > 1)
```

* status
```{r}
#--- status ---
projects %>% count(status)
```

* title
```{r}
projects %>% summarise(n = sum(is.na(title)))
projects %>% count(title, sort = T) %>% filter(n > 1)
```

There seems to be many projects with the same title

* start and end date
```{r}
projects %>% 
  mutate(time_diff = interval(startDate, endDate) / duration(num = 1, units = "years")) %>%
  filter(time_diff > 0 & !is.na(time_diff)) %>%
  filter(time_diff < 10) %>%
  ggplot(aes(time_diff)) +
  geom_histogram(binwidth = 0.1)
```

There are 4 rows with `time_diff <= 0` and many rows with either date `NA`. Overall many projects have unrealistic end date far in the future. Let's visualise that

```{r}
projects %>%
  ggplot(aes(endDate)) + geom_histogram()
```
```{r}
projects %>%
  filter(endDate > "2022-01-01") %>%
  ggplot(aes(endDate)) + 
  geom_histogram()
```

I would say that end dates above `2022-01-01` are not trustworthy. Should look into it, because there are many such




#Data exploration
Let's examine the correlation between project duration and the cost. Let's just look at projects that run from 0.1 to 10 years and cost more than 100 units
```{r}
projects %>% 
  mutate(time_diff = interval(startDate, endDate) / duration(num = 1, units = "years")) %>%
  filter(between(time_diff, 0.1, 10)) %>%
  filter(between(cost, 1e2, 1e9)) %>%
  ggplot(aes(time_diff, cost)) +
  geom_point(alpha=0.1) + 
  geom_smooth(method="lm") +
  scale_y_log10() + 
  scale_x_log10()
```

On a log-log scale there seems to be a linear relationship between cost and duration. It is not very surprising.

How does the ratio of contribution to cost vary over time
```{r}
projects %>% 
  filter(ratio < 1) %>%
  ggplot(aes(startDate, ratio)) +
  geom_point(alpha=0.1)
```


Median contribution and cost over the years

```{r}
projects %>%
  mutate(year = format(startDate, "%Y")) %>%
  filter(!is.na(startDate)) %>%
  filter(cost < 1e7) %>%
  group_by(year) %>%
  summarise(cost = median(cost, na.rm=T),
            contr = median(contr, na.rm=T)) %>%
  gather(key = "type", value = "value", cost, contr) %>%
  ggplot(aes(year, value, fill = type)) +
  geom_bar(stat="Identity", position = position_dodge()) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
```

# Mapping the collaboration network at country level
First let's create a data frame for connections by separating the participant countries

```{r}
cnx <- projects %>%
  unnest_tokens(destination, participantCountries, to_lower = FALSE) %>%
  filter(coordinatorCountry != destination) %>% 
  left_join(countries, by = c(coordinatorCountry = "country")) %>%
  left_join(countries, by = c(destination = "country")) %>%
  select(-name.x, -name.y) %>%
  group_by(coordinatorCountry, destination) %>%
  summarise(n= n(),
            lon1 = unique(longitude.x),
            lat1 = unique(latitude.x),
            lon2 = unique(longitude.y),
            lat2 = unique(latitude.y)) %>%
  arrange(desc(n)) %>%
  filter(destination != "F") %>% #Probably FR but this connection already exists
  filter(destination != "YU")#Probably former Yugoslavia but let's just drop it now

```

Now that we have our connections let's visualise them on a graph

```{r}
cnx %>%
  filter(n > 500) %>%
  # filter(coordinatorCountry %in% eu) %>%
  # filter(destination %in% eu) %>%
  graph_from_data_frame() %>%
  ggraph(layout = "fr") +
  geom_edge_link(aes(edge_alpha = n, edge_width=n), show.legend = FALSE) +
  geom_node_point(color = "skyblue", size = 5) +
  geom_node_text(aes(label = name), size = 2.5, color = "black") +
  theme_void()
```

Connections as great circle distances

```{r}
cnx

color2hex <- function(color, alpha_hex = "FF"){
  # Turns a color's R name into its hex representation 
  c <- col2rgb(color)
  sprintf("#%02X%02X%02X%s", c[1], c[2], c[3], toupper(alpha_hex))
}


map("world", col="#292929", fill=TRUE, bg="#000000", lwd=0.05)

#For alpha gradient
pal <- colorRampPalette(c("#111111", "#aaaaaa")) #Does the interpolation
alpha_vals <- str_extract(pal(256), ".{2}$") #Extract 1 byte

#replace the alpha byte with the interpolated values
colors <- str_replace(color2hex("springgreen", alpha="FF"), ".{2}$", alpha_vals)
alpha_factor <- max(cnx$n)

for(i in 1:nrow(cnx)){
  #Interpolate the great circle lines
  inter2 <- gcIntermediate(c(cnx$lon1[i],
                             cnx$lat1[i]),
                           c(cnx$lon2[i],
                             cnx$lat2[i]), n=100,
                           addStartEnd=TRUE,
                           breakAtDateLine = TRUE)
  
  colindex <- round( (cnx$n[i] / alpha_factor) * (length(colors)-1) ) + 1
  if(is.list(inter2)){
    lapply(inter2, function(x) lines(x, col=colors[colindex], lwd=0.3))
  }else{
    lines(inter2, col=colors[colindex], lwd=0.3)
  }
  lines(c(cnx$lon1[i], cnx$lon2[i]),
        c(cnx$lat1[i], cnx$lat2[i]), col=colors[colindex], lwd=0.3)
}
```

